Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity

### 2017-06-02 manuscript.Rmd
add text, fix typos, polish things up

---
#author: Fran\c{c}ois Michonneau
title: "Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity"
date: '`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    #includes:
    #  before_body: ../template/doc_prefix.tex
    #  in_header: ../template/preamble.tex
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::word_document2: null
fontsize: 12pt
link-citations: yes
csl: nature.csl
bibliography: /home/francois/Library/dark-plankton.bib
---

@@

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood, and modeling the impact of human activities is limited to a few taxonomic groups. A major hinderance is that we do not have enough information about which species live where, and an unknown but potentially significant portion of marine biodiversity remains to be discovered and described. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016]. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functionning are irreversibly altered.

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. These factors have contributed to the lack of synthesis and have made the identification of knowledge gaps difficult. However, with the development of public databases for natural history records (e.g. iDigBio [@Page2015]), species occurrences (e.g., GBIF [@Edwards2004], OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), we can now evaluate what we know. Here, we aggregate publicly available data for marine invertebrates that occur in the United States, in the Gulf of Mexico, and in the Pacific North-West, to highlight gaps in our knowledge. We show that public databases are relatively incomplete, and that most species are known from a few speciemens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but we also need more comprehensive data on species distributions and abundances. This can only be achieved with additional sampling and a commitment to training more taxonomists and experts that can identify the material collected.

<!-- Species identification from DNA sequences is poised to revolutionize our understanding of biodiversity, to facilitate environmental monitoring, and to disentangle trophic networks. However, this approach relies on comprehensive databases of accurately identified sequences.  -->

@@

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio and OBIS databases for all records they included in the same geographic boundaries as the lists. After filtering these lists to only include the same higher taxonomic level taxa as the species lists, and keeping only records for which the species could be match to a species name in WoRMS, we had for the Gulf of Mexico `r nrow(idigbio_gom_records)` and `r nrow(obis_gom_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_gom_records$worms_valid_name)` and `r dplyr::n_distinct(obis_gom_records$worms_valid_name)` species. For the Pacific Northwest, we retrieved `r nrow(idigbio_kozloff_records)` and `r nrow(obis_kozloff_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_kozloff_records$worms_valid_name)` and `r dplyr::n_distinct(obis_kozloff_records$worms_valid_name)` species. For the Gulf of Mexico, the majority of the species included in the expert lists are recorded in both iDigBio and OBIS (`r format_output(species_in_common$n_spp_gom_idigbio_obis/species_in_common$n_spp_gom*100)` %), or OBIS (`r format_output(species_in_common$n_spp_gom_obis/species_in_common$n_spp_gom*100)` %) (Fig. \@ref(fig:database-overlap)). However, for the Pacific Northwest, only `r format_output(species_in_common$n_spp_kozloff_idigbio_obis/species_in_common$n_spp_kozloff*100)`% and `r format_output(species_in_common$n_spp_kozloff_obis/species_in_common$n_spp_kozloff*100)`% of the species from the regional list had records in OBIS and iDigBio or OBIS respectively (Fig. 1).

@@

Species that are in the database but not in the list could be: (1) records not considered by the authors of the checklist: either because they were not accessible or known to the authors; (2) records erronously excluded the records (2) records that have been discovered since the list has been assembled; (3) records with incorrect identification; (4) records with incorrect geographic coordinates: their associated geographic information in the database places them in the Gulf of Mexico or the Pacific Northwest, but they were actually collected elsewhere.

@@

These results suggest that even when considering species lists that have been curated and that represent possibly the most accurate knowledge for this geographic area, today we could only reconstruct these lists partially based on digitized information. For instance, based on OBIS data the list would be `r format_output(summary_richness_per_db_gom$prop_obis_in_gom[summary_richness_per_db_gom$taxon_name == "Porifera"]*100)`% complete for sponges to `r format_output(summary_richness_per_db_gom$prop_obis_in_gom[summary_richness_per_db_gom$taxon_name == "Mollusca"]*100)`% for molluscs. However, these levels drop significantly if we consider species for which museum specimens are available from  `r format_output(summary_richness_per_db_gom$prop_idigbio_in_gom[summary_richness_per_db_gom$taxon_name == "Mollusca"]*100)`% for molluscs to `r format_output(summary_richness_per_db_gom$prop_idigbio_in_gom[summary_richness_per_db_gom$taxon_name == "Annelida"]*100)`% for annelids. Similar patterns can be observed for the Pacific Northwest, but overall the levels of data availability are lower: from `r format_output(summary_richness_per_db_koz$prop_obis_in_pnw[summary_richness_per_db_koz$taxon_name == "Porifera"]*100)`% for sponges to `r format_output(summary_richness_per_db_koz$prop_obis_in_pnw[summary_richness_per_db_koz$taxon_name == "Annelida"]*100)`% for annelids for the proportion of species that have records in OBIS, and from  `r format_output(summary_richness_per_db_koz$prop_idigbio_in_pnw[summary_richness_per_db_koz$taxon_name == "Porifera"]*100)`% for sponges to `r format_output(summary_richness_per_db_koz$prop_idigbio_in_pnw[summary_richness_per_db_koz$taxon_name == "Mollusca"]*100)`% for molluscs for the proportion of species that have records in iDigBio. Museum specimens are critical to validate the occurence of species [@Rocha2014], especially for marine invertebrates for which species limits based on morphology have been known to be challenged by other lines of evidence including molecular data [@Knowlton2000]. Some of the discrepencies observed between the number of species for which occurence data and museum specimens are available suggest that additional digitization of collections (including capturing their geographic information) are needed.

@@

```{r map-diversity-sampling, fig.cap='Spatial distribution of the number of samples (A) and species richness (B) for marine invertebrates based on specimen data available from iDigBio within the limits of the Exclusive Economic Zone for the contiental USA. (\\#fig:map-diversity-sampling)'}
knitr::include_graphics("figures/map_diversity_sampling.pdf")
```

These patterns of spatial distribution of the species richness are strongly confounded by the sampling effort. When investigating the relation between the number of samples and the number of species for each cell of the map (Fig. \@ref(fig:correlation-sampling)), we can see that until reaching about 100 records per cell there is little redundancy in the sampled species.

@@

The inadequacy of the sampling effort compared to the diversity of the fauna is also reflected by the fact that `r prop_singleton_species`% of the species are known from a single record. The strong correlation between the number of samples collected and the species richness suggests that any new sampling effort will lead to new records (Fig. \@ref(fig:map-diversity-sampling)) and new species. Another indication of this inadequacy is the relation between the number of collected specimen and the number of species recorded (Fig. \@ref(fig:idigbio-through-time)). Initial explorations in the late ninteeth century lead to an initial exponential growth in the knowledge of the fauna, that tapered off, until the 1970's when intensive collecting efforts led to an increase in the recording of new taxa (Fig. \@ref(fig:idigbio-through-time)). Since then, the slowing down of collections in the 1970 has been accompanied with a slowing down in the discovery of new records: `r n_spp_comparison$n_spp_2017_2000` species have been documented for the first time since 2000 (`r format_output(n_spp_comparison$n_spp_2017_2000/18)` species.year^-1^), while `r n_spp_comparison$n_spp_1970_1960` have been documented between 1960 and 1970 (`r format_output(n_spp_comparison$n_spp_1970_1960/11)` species.year^-1^). Coordinated large-scale biodiversity expeditions can produce a wealth of data and new records. For instance, a recent study in Western Australia reported that 135 samples led to 805 species of crustaceans and polychates, 94.6% and 72% of which were undescribed [@Poore2014].

@@

Modelling the geographic distributions of species, and how they can be impacted by global change has undergone important advances recently. However, it is generally recognized that to obtain reliable results such efforts require to have accurate geographic information for a minimum of 10 specimens but having at least 30 is preferred. Based on iDigBio data, `r prop_10_specimens_species`% of species have less than 10 specimens and `r prop_30_specimens_species`% have less than 30 specimens. This suggests that our understanding of the  geographic range of species is limited as most species are known from only a few locations, while efforts to estimate their current and future ranges based on occurrence data are rendered moot by the low number of data points available.

There are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups represent the bulk of the records for marine invertebrates in museum collections, the small organisms, including representant of the meiofauna are virtually absent. Of the `r n_idigbio_phyla` phyla occurring in iDigBio, `r records_rare_phyla$n_rare_phyla` are represented by less than 500 records, and make up `r records_rare_phyla$prop_rare_phyla`% of the total number of records.

@@

Since its introduction in the early 2000's, DNA barcoding, the sequencing of a common locus across taxa bearing species-specific information, has revitalized taxonomy by facilitating species delineation [@Meyer2005-plos]. More recently, advances in high-throughput sequencing technologies have enabled the parallel sequencing of thousands of species that occur in environmental samples [@Leray2015]. This approach opens up new research venues it becomes possible to estimate the diversity from species-rich samples that were virtually inaccessible before. This is particularly true in marine communities that include multiple phyla and organisms varying in size by order of magnitude, making identification challenging. However, the utility of the data generated by these projects is enhanced by the possibility of matching the DNA sequences obtained to previously sequenced and correctly identified species [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies may soon be used routinely to monitory biodiversity [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases.

@@

The proportion of species that have DNA barcodes available is relatively constant across the better studied phyla at the scale of the EEZ. It varies between 12% for bryozoans to 28% for cnidarians. The proportions for species known to occur in the Gulf of Mexico are usually similar than for the US, while typically a higher fraction of species from the Pacific Northwest are available. Geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records)` BOLD records analyzed here have no geographic coordinates information associated with them), and many marine invertebrate species are known to be cryptic complexes that are not as widely distributed as previously thought. Therefore specimens identified under the same species but sampled in distant locations, could represent cryptic species. It is thus crucial to ensure that accurate geographic information is provided when DNA barcodes are published, and that additional samples are collected, to capture more accurately their geographic range and the extent of their genetic diversity. This high level of cryptic diversity is reflected in the fact that `r stat_bold$n_spp_muti_bin` species out of `r stat_bold$n_spp` include more than 1 BIN. This number is confounded by some taxonomic identification issues as `r stat_bold$n_shared_bins` BINs are shared across more than one species.

@@

## What to do next?

* Improve data quality:
  - easy: fix mispelling in identifications
  - improve identification of unidentified lots (requires taxonomists/specialists to be able to identify the specimens)
  - add and validate geographic information associated with specimens
  - develop infrastructure and tools to continuously check and validate data
    quality in databases.

However, taxonomic expertise is dwindling, and there are now only a handful of US natural history museums with active marine invertebrates collections. Museum collections have been recording our knowledge about species and their geographic ranges, but until recently, this information was fragmented across institutions and difficult to synthesize. With collection digitization and other large-scale projects have aimed at gathering biodiversity information, public databases can now be used to enable synthesis and highlight knowledge gaps. Here we use these data to synthetic the state of knowledge for marine invertebrates, and demonstrate that the stage is set to combine these data sources to inform where future efforts for documenting biodiversity.

Based on these results, we recommend that additional field expeditions are organized to sample more thoroughly the marine environment. The data from museum collections make clear that any additional samples will contain new species and new records. However, taxonomists are needed to identify and describe these new species. The samples collected during these expeditions need to be photographed to document live appearance that would otherwise be lost during preservation. We also need to capture metadata associated with the specimens to be able to understand better the physical characteristics of the habitats for these species. The specimens need to be sub-sampled and have tissues suitable for genomic analysis preserved, so we can sequence their DNA barcode, or other genomic resources.

; and we estimate how many of the species have genetic data publicly available that would be useful in the context of identifying samples from metabarcoding studies

Using metabarcoding is a promising approach to screen an habitat for its potential for species that are not found in public database, to maximize the usefulness in the future of the data generated by an expedition. Metabarcoding approaches need to complement and take advantage of the DNA barcodes already available. Additionally, in the future, data generated by metabarcoding efforts can be re-interpreted as more DNA barcodes become available. Building the library of DNA barcodes in parallel of metabarcoding efforts also ensures that matches in the database will correspond to actual specimens deposited in natural history collections; ruling out possible artifacts of the method. The knowledge will also build on existing natural history data rather than having to restart from scratch. Beyond their utility in documenting diversity, metabarcoding is also promising for understanding the functioning of ecosystems.

## Conclusions

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems. Additionally,

Biodiversity discovery and documentation has been transformed in the 21st century by molecular data to facilitate species identification, delineate species. And more recently, the development of high throughput sequencing technologies is used to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species, and enabled collaborative research programs such as the Census of Marine Life. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. Despite early criticisms and now well-understood limitations, DNA barcoding has been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification: sequences with no match make it easier to flag specimens that warrant further taxonomic attention.

@@

@@

![Proportion of species occurring in natural history collections that have been DNA barcoded](figures/proportion_barcoded.pdf)

@@

Except for the few phyla that are known to include only a small number of species (Cycliophora, Phoronida, Chaetognatha), microscopic groups, typically component of the meiofauna, are disproportionally under-represented in museum collections, and none or few species are available in public databases.

Our results highlight that most of our knowledge of marine biodiversity come from large expeditions carried in the late 19th century and after World War II. Since then, sampling effort has dramatically decreased, and taxonomic expertise is waning. Yet, our ability to assign DNA sequences to species names will greatly improve the value of the data generated by meta-barcoding studies.

and reflect that a large fraction of the diversity of marine invertebrates remain to be sequenced.

@@

Marine invertebrates are however disproportionally under-represented in DNA sequence databses. And they have a higher rate of misidentification. There are however some pitfalls to this number. It might actually be even higher given that they have been identified by a single taxonomist given the lack of redundancy in expertise.

@@

We also sequenced xx larvae occurring in the plankton at the Whitney Marine Lab, representing yy species (estimated from genetic distances). Of these yy species, only zz% had good matches in BOLD/GenBank. In both cases, taxa1, taxa2, and taxa3 are the most underrepresented in BOLD.

@@

Total number of collections in the US, how many have marine invertebrate collections, how many have active curators.

>>>>

---
#author: Fran\c{c}ois Michonneau
title: "Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity"
date: '`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    #includes:
    #  before_body: ../template/doc_prefix.tex
    #  in_header: ../template/preamble.tex
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: no
fontsize: 12pt
link-citations: yes
csl: nature.csl
bibliography: /home/francois/Library/dark-plankton.bib
---

@@

Museum collections have been recording our knowledge about species and their geographic ranges, but until recently, this information was fragmented across institutions and difficult to synthesize. The digitization of natural history collections and other biodiversity information systems can now be used to enable synthesis and highlight knowledge gaps. Here we use these data to synthesize the state of our knowledge for marine invertebrates in the continental US. We show that current knowledge and sampling efforts are inadequate given the diversity of the fauna and the geographic extent of US. While recent high-throughput sequencing technologies promise to greatly speed up the documentation of marine species and their ranges, its value would be greatly improved by anchoring the results into existing DNA sequence data and associated voucher. However, the best documented phyla have at most 25% of their species sequenced for COI. More sampling, coordinated at the scale of the US, is needed to improve our knowledge of species and their geographic ranges. This data is needed to detect the impact of global change and other perturbations. Advances in computing, information systems, and sequencing technologies, have set the stage to inform where future efforts for documenting biodiversity should be applied.

## Introduction

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood, and modeling the impact of human activities is limited to a few taxonomic groups. A major hindrance is that we do not have enough information about which species live where, and an unknown but potentially significant portion of marine biodiversity remains to be discovered and described. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016]. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functioning is irreversibly altered.

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. These factors have contributed to the lack of synthesis and have made the identification of knowledge gaps difficult. However, with the development of public databases for natural history records (e.g. iDigBio [@Page2015]), species occurrences (e.g., GBIF [@Edwards2004], OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), we can better evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur in the continental United States, the Gulf of Mexico, and the Pacific Northwest, to highlight gaps in our knowledge. We show that public databases are relatively incomplete, and that most species are known from a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but we also need more comprehensive data on species distributions and abundances. This can only be achieved with additional sampling and a commitment to training more taxonomists and experts that can identify the material collected.

@@

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio and OBIS databases for all records they included in the same geographic boundaries as the lists. After filtering these lists to only include the same higher taxonomic level taxa as the species lists, and keeping only records for which the species could be matched to a species name in WoRMS, we had for the Gulf of Mexico `r nrow(idigbio_gom_records)` and `r nrow(obis_gom_records)` records for iDigBio and OBIS respectively, representing `r dplyr::n_distinct(idigbio_gom_records$worms_valid_name)` and `r dplyr::n_distinct(obis_gom_records$worms_valid_name)` species. For the Pacific Northwest, we retrieved `r nrow(idigbio_kozloff_records)` and `r nrow(obis_kozloff_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_kozloff_records$worms_valid_name)` and `r dplyr::n_distinct(obis_kozloff_records$worms_valid_name)` species. For the Gulf of Mexico, the majority of the species included in the expert lists are recorded in both iDigBio and OBIS (`r format_output(species_in_common$n_spp_gom_idigbio_obis/species_in_common$n_spp_gom*100)` %), or OBIS (`r format_output(species_in_common$n_spp_gom_obis/species_in_common$n_spp_gom*100)` %) (Fig. \@ref(fig:database-overlap)). However, for the Pacific Northwest, only `r format_output(species_in_common$n_spp_kozloff_idigbio_obis/species_in_common$n_spp_kozloff*100)`% and `r format_output(species_in_common$n_spp_kozloff_obis/species_in_common$n_spp_kozloff*100)`% of the species from the regional list had records in OBIS and iDigBio or OBIS respectively (Fig. 1).

@@

Species that are in the database but not in the list could be: (1) records not considered by the authors of the checklist: either because they were not accessible or known to the authors; (2) records erroneously excluded the records (2) records that have been discovered since the list has been assembled; (3) records with incorrect identification; (4) records with incorrect geographic coordinates: their associated geographic information in the database places them in the Gulf of Mexico or the Pacific Northwest, but they were actually collected elsewhere.

@@

These results suggest that even when considering species lists that have been curated and that represent possibly the most accurate knowledge for this geographic area, today we could only reconstruct these lists partially based on digitized information. For instance, based on OBIS data the list would be `r format_output(summary_richness_per_db_gom$prop_obis_in_gom[summary_richness_per_db_gom$taxon_name == "Porifera"]*100)`% complete for sponges to `r format_output(summary_richness_per_db_gom$prop_obis_in_gom[summary_richness_per_db_gom$taxon_name == "Mollusca"]*100)`% for molluscs. However, these levels drop significantly if we consider species for which museum specimens are available from  `r format_output(summary_richness_per_db_gom$prop_idigbio_in_gom[summary_richness_per_db_gom$taxon_name == "Mollusca"]*100)`% for molluscs to `r format_output(summary_richness_per_db_gom$prop_idigbio_in_gom[summary_richness_per_db_gom$taxon_name == "Annelida"]*100)`% for annelids. Similar patterns can be observed for the Pacific Northwest, but overall the levels of data availability are lower: from `r format_output(summary_richness_per_db_koz$prop_obis_in_pnw[summary_richness_per_db_koz$taxon_name == "Porifera"]*100)`% for sponges to `r format_output(summary_richness_per_db_koz$prop_obis_in_pnw[summary_richness_per_db_koz$taxon_name == "Annelida"]*100)`% for annelids for the proportion of species that have records in OBIS, and from  `r format_output(summary_richness_per_db_koz$prop_idigbio_in_pnw[summary_richness_per_db_koz$taxon_name == "Porifera"]*100)`% for sponges to `r format_output(summary_richness_per_db_koz$prop_idigbio_in_pnw[summary_richness_per_db_koz$taxon_name == "Mollusca"]*100)`% for molluscs for the proportion of species that have records in iDigBio.

Museum specimens are critical to validate the occurrence of species [@Rocha2014], especially for marine invertebrates for which species limits based on morphology have been known to be challenged by other lines of evidence including molecular data [@Knowlton2000]. Some of the discrepancies observed between the number of species for which occurrence data and museum specimens are available suggest that additional digitization of collections (including capturing their geographic information) are needed.

@@

```{r map-diversity-sampling, fig.cap='Spatial distribution of the number of samples (A) and species richness (B) for marine invertebrates based on specimen data available from iDigBio within the limits of the Exclusive Economic Zone for the continental USA. (\\#fig:map-diversity-sampling)'}
knitr::include_graphics("figures/map_diversity_sampling.pdf")
```

These patterns of spatial distribution of the species richness are strongly confounded by the sampling effort. When investigating the relation between the number of samples and the number of species for each cell of the map (Fig. \@ref(fig:correlation-sampling)), we can see that until reaching about 100 records per cell there is little redundancy in the sampled species (each record represent a different species).

@@

The inadequacy of the sampling effort compared to the diversity of the fauna is also reflected by the fact that `r prop_singleton_species`% of the species are known from a single record. The strong correlation between the number of samples collected and the species richness suggests that any new sampling effort will lead to new observations (Fig. \@ref(fig:map-diversity-sampling)) and new species. Another indication of this inadequacy is the relation between the number of collected specimen and the number of species recorded through time (Fig. \@ref(fig:idigbio-through-time)). Initial explorations in the late ninetieth century lead to an initial exponential growth in the knowledge of the fauna, that tapered off, until the 1970's when intensive collecting efforts led to an increase in the recording of new taxa (Fig. \@ref(fig:idigbio-through-time)). Since then, the slowing down of collections in the 1970 has been accompanied with a slowing down in the discovery of new records: `r n_spp_comparison$n_spp_2017_2000` species have been documented for the first time since 2000 (`r format_output(n_spp_comparison$n_spp_2017_2000/18)` species.year^-1^), while `r n_spp_comparison$n_spp_1970_1960` have been documented between 1960 and 1970 (`r format_output(n_spp_comparison$n_spp_1970_1960/11)` species.year^-1^). Coordinated large-scale biodiversity expeditions can produce a wealth of data and new records. For instance, a recent study in Western Australia reported that 135 samples led to 805 species of crustaceans and polychaetes, 94.6% and 72% of which were undescribed [@Poore2014].

@@

Modeling the geographic distributions of species, and how they can be impacted by global change has undergone important advances recently. However, it is generally recognized that to obtain reliable results such efforts require to have accurate geographic information for a minimum of 10 specimens but having at least 30 is preferred. Based on iDigBio data, `r prop_10_specimens_species`% of species have less than 10 specimens and `r prop_30_specimens_species`% have less than 30 specimens. This suggests that our understanding of the geographic range of species is limited as most species are known from only a few locations, while efforts to estimate their current and future ranges based on occurrence data are rendered moot by the low number of data points available.

There are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups represent the bulk of the records for marine invertebrates in museum collections, the small organisms, including representatives of the meiofauna are virtually absent. Of the `r n_idigbio_phyla` phyla occurring in iDigBio, `r records_rare_phyla$n_rare_phyla` are represented by less than 500 records, and make up `r records_rare_phyla$prop_rare_phyla`% of the total number of records.

@@

Since its introduction in the early 2000's, DNA barcoding, the sequencing of a common locus across taxa bearing species-specific information, has revitalized taxonomy by facilitating species delineation [@Meyer2005-plos]. More recently, advances in high-throughput sequencing technologies have enabled the parallel sequencing of thousands of species that occur in environmental samples [@Leray2015]. This approach opens up new research venues as it becomes possible to estimate the diversity from species-rich samples that were virtually inaccessible before. This is particularly true in marine communities that include multiple phyla and organisms varying in size by order of magnitude, making identification challenging. However, the utility of the data generated by these projects is enhanced by the possibility of matching DNA sequences obtained to previously sequenced and correctly identified species [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies may soon be used routinely to monitor biodiversity [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases.

@@

The proportion of species that have DNA barcodes available is relatively constant across the better studied phyla at the scale of the EEZ. It varies between 12% for bryozoans to 28% for cnidarians. The proportions for species known to occur in the Gulf of Mexico are usually similar than for the US, while typically a higher fraction of species from the Pacific Northwest are available. Geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records)` BOLD records analyzed here have no geographic coordinates information associated with them), and many marine invertebrate species are known to be cryptic complexes that are not as widely distributed as previously thought. Therefore specimens identified under the same species but sampled in distant locations, could represent cryptic species. It is thus crucial to ensure that accurate geographic information is provided when DNA barcodes are published, and that additional samples are collected, to capture more accurately their geographic range and the extent of their genetic diversity. This cryptic diversity is reflected in the fact that `r stat_bold$n_spp_multi_bin` species out of `r stat_bold$n_spp` include more than 1 BIN (BOLD Index Number, operational taxonomic units that closely resemble species [@Ratnasingham2013]).

@@

## Where do we go from there?

Based on the results of this study, we recommend a multi-level approach to not only improve the quality of the existing data, but also to efficiently collect new data.

### Improve existing data

With data aggregation of natural history collections through initiatives like iDigBio, and the compilation of a nearly complete register of marine species with WoRMS, in combination with the development of algorithms for text manipulation, it is increasingly easier to validate and spell check the scientific names of the identifications of museum specimens. Additionally, because WoRMS links the history of the names and their synonymies, it is possible to dissociate the concepts of species identification with the concept of species identity, allowing to compare specimens identified at different time periods.

Similarly, a significant fraction of museum records do not have or have inaccurate geographic information associated with them. While tools have been developed to convert names of locales into GPS coordinates on land, similar approaches may not work for locales at sea. However, geographic information is available in other forms, and can be digitized.

For both the spelling of scientific names, and the geographic information, computing tools can be developed to detect issues and propose solutions. There are opportunities for services using publicly available data to be developed to make curators and collection managers aware of these issues so that they can fixed at the source. It however requires that resources are available to museum collections to address these issues.

Increasing the computing literacy of the staff of museum in order for them to manipulate and use the data from their collection database is equally critical. As data across institutions get pooled into the large biodiversity information systems, it is crucial that data standards such as Darwin Core are embraced by the community.

<!-- - develop infrastructure and tools to continuously check and validate data quality in databases. -->

### Collect better data faster

This study highlights that we have too little data about too few species of marine invertebrates, we recommend that additional coordinated field expeditions are organized to sample more thoroughly the marine environment. Data from museum collections make clear that any additional samples will contain new species and new records. Samples collected during these expeditions need to be photographed to document live appearance that would otherwise be lost during preservation. We also need to capture metadata associated with the specimens to be able to understand better the physical characteristics of the habitats for these species. Specimens need to be sub-sampled and have tissues suitable for genomic analysis preserved for their DNA barcode, or any other genomic resources. Some of these parts of data and specimen collection can be automated and captured digitally directly in the field speeding up data collection.

Coupling these more traditional approaches of biodiversity documentation with metabarcoding is a promising approach to screen an habitat for its potential for species that do not yet occur in public databases. Metabarcoding can complement and take advantage of the DNA barcodes already available. Additionally, in the future, data generated by metabarcoding efforts can be re-interpreted as more DNA barcodes associated with correctly identified specimens become available. Building the library of DNA barcodes in parallel of metabarcoding efforts ensures that matches in the database will correspond to actual specimens deposited in natural history collections; ruling out possible artifacts of the method. Information about species identity and distribution will thus build on existing natural history data rather than being limited to matching DNA sequences across studies.

<!-- Beyond their utility in documenting diversity, metabarcoding is also promising for understanding the functioning of ecosystems. -->

However, taxonomists and experts are needed to collect, sort, process, and identify these samples and the data that come along with them. Even if technological advances can greatly speed up the documentation of species and their distributions, trained personnel is needed. Taxonomic expertise is dwindling, and there are now only a handful of US natural history museums with active marine invertebrates collections.

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems.

<!-- Biodiversity discovery and documentation has been transformed in the 21st century by molecular data to facilitate species identification, delineate species. And more recently, the development of high throughput sequencing technologies is used to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species, and enabled collaborative research programs such as the Census of Marine Life. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. Despite early criticisms and now well-understood limitations, DNA barcoding has been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification: sequences with no match make it easier to flag specimens that warrant further taxonomic attention. -->

@@

<!-- widening gap with other groups (vertebrates, plants) for which more comprehensive data is available -->

@@

@@

@@

@@

@@
